\bf\mftitle

### 2011-01-27 doc/jInfer/architecture/architecture.tex
fixed overfull hboxes by package names
moved factory pattern to arch from twostep

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\emph{#1}}

@@

\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\

@@

Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
Each \code{Regexp<T>} instance has one of the enum \code{RegexpType} type:
\begin{itemize}
	\item Lambda ($\lambda$) - empty string (also called $\epsilon$ in literature),
	\item Token - a letter of the alphabet,
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$,
	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$,
	\item Permutation - shortcut for all possible permutations of regular expressions. Our syntax to write down permutation is $(a\& b\& c\& d)$.
\end{itemize}
Type of regexp is held in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.

@@

Regexp has to obey constraits:
\begin{itemize}
	\item type, children and interval have to be non-null references,
	\item when type is lamba, content and interval has to be null,
	\item when type is token, content has to be non-null,
	\item when type concatenation, alternation or permutation, content has to be null.
\end{itemize}
These constraits are checked by constructors, so the best way to construct new regexps is by using 
methods \code{getToken(), getConcatenation()} etc.

@@

Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full}.
In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{AbstractStructuralNode} variables. And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.

@@

\subsection{Programmatic view}
From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
\begin{itemize}
	\item \code{IGGenerator}
	\item \code{Simplifier}
	\item \code{SchemaGenerator}
\end{itemize}
A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{SimplifierImpl}.
Annotation required for the framework to recognize such a class as an inference module is the following
\begin{verbatim}
	@ServiceProvider(service = <interface>.class)
\end{verbatim}
for example
\begin{verbatim}
	@ServiceProvider(service = Simplifier.class)
\end{verbatim}
The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is to be executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameters now in more detail.

@@

For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCapabilities()}. This method returns a list of strings representing the abilities, or "capabilities" of this module.
An inference module can query the capabilities of the \emph{next} module in the inference chain by invoking the 
\begin{verbatim}
RunningProject.getNextModuleCaps()
\end{verbatim}
method to obtain the next module encapsulated in the \code{Capabilities} interface.

@@

At this point it is important to understand that when using lookups this way, all service providers are kept as singletons. In case a provider needs to change its state during providing the service (and needs to be ``fresh" each time it is retrieved), it is best to convert this to a factory design pattern: the service provider registered is just a factory returning instances that do the actual work.

@@

>>>>

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\emph{#1}}

@@

\noindent\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\

@@

Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{Reg\.exp\.Type}.
Each \code{Regexp<T>} instance has one of the enum \code{RegexpType} type:
\begin{itemize}
	\item Lambda ($\lambda$) - empty string (also called $\epsilon$ in literature),
	\item Token - a letter of the alphabet,
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$,
	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$,
	\item Permutation - shortcut for all possible permutations of regular expressions. Our syntax to write down permutation is $(a\& b\& c\& d)$.
\end{itemize}
Type of regexp is in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLa\.mb\.da()}, \code{isTo\.ken()}, \code{isCon\.ca\.te\.na\.tion()}, \code{isAlt\.er\.na\.tion()}, \code{isPer\.mu\.ta\.tion()}.

@@

Regexp has to obey constraits:
\begin{itemize}
	\item type, children and interval have to be non-null references,
	\item when type is lamba, content and interval has to be null,
	\item when type is token, content has to be non-null,
	\item when type concatenation, alternation or permutation, content has to be null.
\end{itemize}
These constraits are checked by constructors, so the best way to construct new regexps is by using 
methods \code{get\.To\.ken(), get\.Con\.ca\.te\.na\.tion()} etc.

@@

Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full2}.
In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{Abstr\.act\.Stru\.ctu\.ral\.Node} variables. And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.

@@

\subsection{Programmatic view}
From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
\begin{itemize}
	\item \code{IGGenerator}
	\item \code{Simplifier}
	\item \code{SchemaGenerator}
\end{itemize}
A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{Sim\.pli\.fier\.Impl}.
Annotation required for the framework to recognize such a class as an inference module is the following
\begin{verbatim}
	@ServiceProvider(service = <interface>.class)
\end{verbatim}
for example
\begin{verbatim}
	@ServiceProvider(service = Simplifier.class)
\end{verbatim}
The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is to be executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameters now in more detail.

@@

For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCa\.pa\.bi\.li\.ties()}. This method returns a list of strings representing the abilities, or "capabilities" of this module.
An inference module can query the capabilities of the \emph{next} module in the inference chain by invoking the 
\begin{verbatim}
RunningProject.getNextModuleCaps()
\end{verbatim}
method to obtain the next module encapsulated in the \code{Capabilities} interface.

@@

@@

\subsection{Factory pattern}
\begin{figure}
  \centering
	\subfigure[Same instance of class returned by NetBeans in two successive inference runs]
	{
	  \centering\includegraphics[scale=\myscale]{factory_pattern1}
    \label{factory_pattern1}
  }
	\subfigure[Factory solution of singleton lookup class problem]
	{
	  \centering\includegraphics[scale=\myscale]{factory_pattern2}
    \label{factory_pattern2}
  }
\end{figure}
At this point it is important to understand that when using lookups this way, all service providers are kept as singletons. 
We ilustrate this behaviour on \ref{factory_pattern1}, user runs inference first time, classes are looked up and used.
When user clicks run button again, and same instances of classes are returned by lookups.

In case a provider needs to change its state during providing the service (and needs to be ``fresh'' each time it is retrieved), there are two solutions:
\begin{itemize}
	\item force each service providing class (submodule), to implement
some sort of cleanup method, that would restart it into fresh state and make it ready to use by another 
inference run,
	\item or create new instances of stateful classes in each inference run and make service providing classes (those which NetBeans return by lookups) factories.
\end{itemize}
We decided for the latter approach (as ilustrated on fig. \ref{factory_pattern2}
It enables us not only to produce fresh class instance in each inference run, but also factory classes implement obligatory
module methods such as \code{getName}, \code{getDescription} and so on, which are really same in each inference run.
Each submodule is then defined by (at least) two interfaces.
One is the factory interface, like this one (example from \jmodule{TwoStepSimplifier}:
\begin{verbatim}
public interface AutomatonSimplifierFactory extends 
                     NamedModule, Capabilities, UserModuleDescription {
  <T> AutomatonSimplifier<T> create();
}
\end{verbatim}
Second is worker interface, in the example it is \code{AutomatonSimplifier<T>} interface.

This pattern is used thorough jInfer modules to divide module into submodules.
